---
title: "Coursera stats inference"
author: "Will S."
date: "4/2/2021"
output:
  html_document: default
  pdf_document: default
  word_document: default
---
# Title: Coursera stats inference, by Will S. 
 
## Overview 
In part 1, I will demostrate 1,000 exponential distribution samples with n = 40 and lambda = 0.2 will converge asymptotically to a normal distribution, following the CLT. In part 2, I will construct a 95% confidence interval to show that the sources of vitamin C may be irrelevant to the growth of tooth, using the R toothgrowth dataset. 
 
## PART1: Simulation. 
Set the relevant properties of the exp distribution.
```{r exp dist property}
lambda = 0.2 #exp dist property
En = c() #an empty vector to store means
Vn = c() # an empty vector to store variances
m = 1000 #simulation times
op = par(mfrow=c(1,1)) #plot
n = 40 #sample size
```
 
simulation: 
 
```{r simulation}
#simulation begins
for (i in 1:m){
  
  #sample from exp
  W = rexp(n, lambda)
  
  #standardize
  En[i] = (mean(W) - 1/lambda)/((1/lambda)/sqrt(n))
  Vn[i] = sd(W)^2
}
```
 
plot: 
 
```{r plot}
#plot
head(En) #dist mean = 1/lambda = 5
head(Vn) #dist var = (1/lambda)^2 = 25
hist(En, probability = TRUE, breaks = seq(-5,5,by=.5), main = "1,000 samples from means of rexp(40, .2)")
curve(dnorm(x, mean = 0, sd = 1), add= TRUE, col = "red")
```

