---
title: "toothgrowth"
author: "Will S."
date: "4/2/2021"
output:
  pdf_document: default
  html_document: default
---

## PART2: Confidence interval 

Load the data: 

```{r loading data}
data <- ToothGrowth
summary(data)
head(data)
str(data)
library(dplyr)
```

Some exploratory analysis. Two boxplots will display to show the the length of tooth growth group_by-ed variable supp and variable dose. 

```{r exp analysis plot}
#exploratory plotting
with(data, boxplot(len~supp))
with(data, boxplot(len~dose))
```

datavc and dataoj are vectors storing growth values from respective supp. 
 
```{r}
datavc <- data[1:30,1]
dataoj <- data[31:60,1]
```

Moments(Mean, sd, and variance) of length of tooth growth group_by-ed variable supp and variable dose. 

```{r moments}
#moments by supp
datasupp <- data %>%
  group_by(supp) %>%
  summarize(len_mu = mean(len), len_sd = sd(len), len_var = var(len))
datasupp

#moments by dose
datadose <- data %>%
  group_by(dose) %>%
  summarize(len_mu = mean(len), len_sd = sd(len), len_var = var(len))
datadose
```

We construct a 95% conf interval. The results from directly calculation will be compared with that using t.test function$conf. Both show that the intercal contain 0, meaning we cannot eliminate the possibility of the fact that VC is not necessarily better than OJ in terms of providing vitaminC supplement.

```{r conf interval}
#by supp, can be treated as 2 independent samples, each 30
#VC-OJ
s <- sqrt((29*(as.numeric(datasupp[1,4]) + 
                 as.numeric(datasupp[2,4])))/58)
intvl <- as.numeric(datasupp[2,2] - datasupp[1,2]) + c(-1,1)*qt(0.975, 58)*s*sqrt(1/15)
t_conf <- t.test(datavc, dataoj)$conf
```

```{r intvl}
intvl
```

```{r t_conf}
t_conf
```


